---
title: "Figures S6A, S6B, S6C, and S6D"
author: "Asaf Shilo/Gianluca Pegoraro"
date: "August 31st 2022"
output: github_document
---

### Read and Process the Object Level Data

Load required packages.
```{r loadAndClean}
library(plyr)
library(tidyverse)
library(data.table)
library(stringr)
library(ggthemes)
library(ggpubr)
library(gghalves)
library(rstatix)
library(FSA)
library(curl)
```

Set output folder.

```{r, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  fig.path = 'output/',
  message = FALSE,
  warning = FALSE
  )
```

Set the palette and the running theme for ggplot2.

```{r setThemeandPalette}
theme_set(theme_minimal())
theme_update(axis.text.x = element_text(
angle = -90,
hjust = 0,
vjust = 0.5
))
```

Read and Change variable types in experimental metadata data.table

```{r readMetadata}
dt_md <- fread("metaData/Layout.txt")
dt_md <- dt_md[probe != "",]
dt_md[, `:=`(probe = factor(probe, levels = c("None", 
                                              "FGFR2_Full",
                                              "FGFR2_IIIb",
                                              "FGFR2_IIIc",
                                              "TBP",
                                              "FGFR2_D7_10",
                                              "FGFR2_D7_10_B3",
                                              "All_IIIb",
                                              "All_IIIc" )))]
```

### Download the data if needed

Download unzip the Columbus results of the experiments from Figshare if they have not been already downloaded.

```{r downlaod-data}
if(!dir.exists("ObjectLevelData")) {
  URL <- "https://figshare.com/ndownloader/files/36988441"
  curl_download(URL, "input.zip")
  unzip("input.zip")
}
```

### Read and process the cell level data

Set RegEx patterns for directory searches for cell level data.

```{r regexFilename}
pat_c <- ".*Nuclei Final\\[0\\]\\.txt$" # Pattern for Single Cell data files
```

Create a list of the RegEx patterns set in the previous chunk. **Important:** the list names will be carried over all the next steps!!!

-   c = cell level data

```{r patList}
pat_list <- list(c= pat_c)
```

Recursively search the `ObjectLevelData` directory and its subdirectories for files whose name includes the RegEx patterns defined two chunks above. The `path_list` functon outputs absolute file names. `path_list` is a list containing all the filenames on a per cell-level.

```{r directorySearch}
list_files <- function(x) {
  dir(
  path = 'ObjectLevelData/',
  pattern = x,
  full.names = TRUE,
  recursive = TRUE,
  include.dirs = TRUE
  )
}

path_list <- llply(pat_list, list_files) 
```

Extract file names from absolut path and set them as list element names. 

```{r trimNames}
trim_names <- function(x) {
  names(x) <-
  basename(x) # This assigns the filename to the file that it is read
  y <- x ## This is necessary because of scoping issues
}

path_list <- llply(path_list, trim_names) 
```

Recursively read and merge object level data files as data.frames. Rows are labeled with relative filenames (The `.id` variable). This and the previous chunks are slightly modified tricks adopted from H. Wickam ["Tidy Data" paper](http://vita.had.co.nz/papers/tidy-data.pdf).

```{r readMerge}
read_merge <- function(x) {
  dt <- as.data.table(ldply(x, fread, sep = "\t", na.strings = NULL))
}

dt_list <- llply(path_list, read_merge)
```

Separate the cell level data from the distance and spot level data.

```{r deleteDtList}
dt_cell <- dt_list$c

rm(dt_list)
```

Change attribute names and create a new `im_acq` variable to indicate 2D or 3D imaging acquisitions.

```{r changeNames}
setnames(
  dt_cell,
  c("Row",
    "Column",
    "ScreenName",
    "WellName",
    
    "Nuclei Final - Nucleus Roundness",
    "Nuclei Final - Number of Positive 488- per Cell",
    "Nuclei Final - Number of Positive 562- per Cell",
    "Nuclei Final - Number of Positive 647- per Cell"),
  c("row",
    "column",
    "plate",
    "well",
   
    "roundness",
    "spot_n_488",
    "spot_n_562",
    "spot_n_647")
  )
```

Join measurementa data with experimental metadata.

```{r dt}
setkey(dt_md, row, column)
setkey(dt_cell, row, column)

dt_full <- dt_cell[dt_md, nomatch = 0] 
```

Generate datasets for plotting.

```{r df}
dt_IIIB <- filter(dt_full, probe == 'FGFR2_IIIb')
dt_IIIC <- filter(dt_full, probe == 'FGFR2_IIIc')
dt_Full <- filter(dt_full, probe == 'FGFR2_Full')
dt_TBP <- filter(dt_full, probe == 'TBP')
dt_D7_10 <-filter(dt_full, probe == 'FGFR2_D7_10')
dt_D7_10_b3 <-filter(dt_full, probe == 'FGFR2_D7_10_B3')
dt_All_B <- filter (dt_full, probe == 'All_IIIb')
dt_All_C <- filter (dt_full, probe == 'All_IIIc')
dt_MP<- filter(dt_full, probe !='FGFR2_Full' & probe !="FGFR2_D7_10")
```

### Plotting

Plotting halves FGFR2-Full in B1-488 System. 

```{r FigS6_A, echo = F}
spot_Full_halves_log <- ggplot(dt_Full, aes(x = cell_line, y = spot_n_488))


spot_Full_halves_log + 
                      geom_half_boxplot(side = "r",nudge = 0.025, outlier.shape = NA, colour="black", fill ="red3") +
                      geom_half_violin(side = "l",width = 1, nudge = 0.025, fill = "red3")+

                scale_y_continuous(trans='log10') + 
                labs(y = "FGFR2-Full HCR Spots per cell") +
                theme_classic()+
                theme(
                legend.position ="none",
                plot.title = element_text(size = 10, hjust = 0.5, face = "bold"),
                axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 10, color="black", face="bold"),
                axis.text = element_text(size=10,color="black", face="bold"))
```


Plotting halves FGFR2-IIIb in B1-488 System

```{r FigS6_B, echo = F}
spot_B1_IIIb_halves <- ggplot(dt_IIIB, aes(x = cell_line, y = spot_n_488))


spot_B1_IIIb_halves + 
                      geom_half_boxplot(side = "r",nudge = 0.025, outlier.shape = NA, colour="black", fill ="green4") +
                      geom_half_violin(side = "l",width = 1, nudge = 0.025, fill = "green4")+

 scale_y_continuous(trans='log10') + 
            labs(y = "FGFR2-IIIb HCR Spots per cell") +
                theme_classic()+
                theme(
                legend.position ="none",
                plot.title = element_text(size = 10, hjust = 0.5, face = "bold"),
                axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 10, color="black", face="bold"),
                axis.text = element_text(size=10,color="black", face="bold"))
```

Plotting halves FGFR2-IIIc in B1-488 System

```{r FigS6_C, echo = F}
spot_IIIc_halves_log <- ggplot(dt_IIIC, aes(x = cell_line, y = spot_n_488))


spot_IIIc_halves_log + 
                      geom_half_boxplot(side = "r",nudge = 0.025, outlier.shape = NA, colour="black", fill ="blue3") +
                      geom_half_violin(side = "l",width = 1, nudge = 0.025, fill = "blue3")+

                scale_y_continuous(trans='log10') + 
                labs(y = "FGFR2-IIIc HCR Spots per cell") +
                theme_classic()+
                theme(
                legend.position ="none",
                plot.title = element_text(size = 10, hjust = 0.5, face = "bold"),
                axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 10, color="black", face="bold"),
                axis.text = element_text(size=10,color="black", face="bold"))
```

Plotting halves TBP in B2-562 System.

```{r FigS6_D, echo = F}
spot_TBP_halves_log <- ggplot(dt_TBP, aes(x = cell_line, y = spot_n_562))


spot_TBP_halves_log +
  geom_half_boxplot(side = "r",nudge = 0.025, outlier.shape = NA, colour="black", fill ="grey70") +
                    geom_half_violin(side = "l",width = 1, nudge = 0.025, fill = "grey70")+
  
                  scale_y_continuous(trans='log10') + 
                  labs(y = "TBP HCR Spots per cell") +
                  theme_classic()+
                  theme(
                  legend.position ="none",
                  plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
                  axis.title.x = element_blank(),
                  axis.text.x = element_text(angle = 45, hjust = 1),
                  axis.title = element_text(size = 10, color="black", face="bold"),
                  axis.text = element_text(size=10,color="black", face="bold"))
```


Document the information about the analysis session

```{r sessionInfo, include=TRUE, echo=TRUE, results='markup'}
sessionInfo()