---
title: "Supp Fig 6"
author: "Asaf Shilo"
date: "march 18th 2022"
output:
  html_document: default
  pdf_document: default
---

### Read and Process the Object Level Data

Load required packages.
```{r loadAndClean}
library(plyr)
library(tidyverse)
library(data.table)
library(stringr)
library(ggthemes)
library(hexbin)
library(ggpubr)
library(viridis)
library(RColorBrewer)

```

```{r, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  fig.path = 'R_Output_Supp.Fig.6/',
  dev = c("png","tiff","pdf"),
  message = FALSE,
  warning = FALSE
  )
```

Set the palette and the running theme for ggplot2.
```{r setThemeandPalette}
theme_set(theme_minimal())
theme_update(axis.text.x = element_text(
angle = -90,
hjust = 0.5,
vjust = 0.5,
size = 20
))
```

Read and Change variable types in experimental metadata data.table
```{r readMetadata}
dt_md <- fread("ExperimentalMetaData/GlobalLayout.txt")
dt_md[, `:=`(treatment = factor(treatment, levels = c("si-Control",
                                              "si-ESRP1-2"
                                                     )))]
```

Set RegEx patterns for directory searches for cell level data.
```{r regexFilename}
pat_c <- ".*Nuclei Final\\[0\\]\\.txt$" # Pattern for Single Cell data files
```

Create a list of the RegEx patterns set in the previous chunk. **Important:** the list names will be carried over all the next steps!!!

* c = cell level data

```{r patList}
pat_list <- list(c= pat_c)
```

Recursively search the `ObjectLevelData` directory and its subdirectories for files whose name includes the RegEx patterns defined two chunks above. The `path_list` functon outputs absolute file names. `path_list` is a list containing all the filenames on a per cell-level.
```{r directorySearch}
list_files <- function(x) {
  dir(
  path = 'ObjectLevelData/',
  pattern = x,
  full.names = TRUE,
  recursive = TRUE,
  include.dirs = TRUE
  )
}

path_list <- llply(pat_list, list_files) 
```

Extract file names from absolut path and set them as list element names. 
```{r trimNames}
trim_names <- function(x) {
  names(x) <-
  basename(x) # This assigns the filename to the file that it is read
  y <- x ## This is necessary because of scoping issues
}

path_list <- llply(path_list, trim_names) 
```

Recursively read and merge object level data files as data.frames. Rows are labeled with relative filenames (The `.id` variable). This and the previous chunks are slightly modified tricks adopted from H. Wickam ["Tidy Data" paper](http://vita.had.co.nz/papers/tidy-data.pdf).
```{r readMerge}
read_merge <- function(x) {
  dt <- as.data.table(ldply(x, fread, sep = "\t", na.strings = NULL))
}

dt_list <- llply(path_list, read_merge)
```

Separate the cell level data from the distance and spot level data.
```{r deleteDtList}
dt_cell <- dt_list$c

rm(dt_list)
```

Change attribute names and create a new `im_acq` variable to indicate 2D or 3D imaging acquisitions.
```{r changeNames}
setnames(
  dt_cell,
  c("Row",
    "Column",
    "ScreenName",
    "WellName",
    "Nuclei Final - Number of Positive 647- per Cell",
    "Nuclei Final - Number of Positive 562- per Cell",
    "Nuclei Final - Number of Positive 488- per Cell"
  ),
  c("row",
    "column",
    "plate",
    "well",
    "Far_Red",
    "Red",
    "Green")
  )
```

Join measurementa data with experimental metadata.
```{r}
setkey(dt_md, row, column)
setkey(dt_cell, row, column)

dt_full <- dt_cell[dt_md, nomatch = 0] 
```


```{r, echo = F}
dt_ESRP <- dt_full %>% filter(probe=="ESRP" )
```


ESRP1
```{r, echo = F}
spot_ESRP1_violin <- ggplot(dt_ESRP, aes(x = treatment, y =  Red))

spot_ESRP1_violin + geom_violin(fill ="skyblue1") +
  geom_boxplot(width=0.1, color="black", fill ="white", outlier.shape = NA)+
           ylim(0,125)+
  labs(y = "ESRP1 HCR Spots per cell") +
  theme_classic()+
  theme(
  legend.position ="none",
                axis.title.x = element_blank(),
                axis.title = element_text(size = 16, color="black", face="bold"),
                axis.text = element_text(size=16,color="black", face="bold"))
        
```


ESRP2
```{r, echo = F}
spot_ESRP2_violin <- ggplot(dt_ESRP, aes(x = treatment, y =  Far_Red))

spot_ESRP2_violin + geom_violin(fill = "skyblue3") +geom_boxplot(width=0.1, color="black", fill ="white", outlier.shape = NA)+
           ylim(0,150)+
  labs(y = "ESRP2 HCR Spots per cell") +
  theme_classic()+
  theme(
  legend.position ="none",
                axis.title.x = element_blank(),
                axis.title = element_text(size = 16, color="black", face="bold"),
                axis.text = element_text(size=16,color="black", face="bold"))
        
```


```{r, fig.length = 14, fig.width = 7}
spot_ESRP1_ESRP2 <- ggplot(dt_ESRP, aes(x = Far_Red, y =  Red))

spot_ESRP1_ESRP2 + geom_point(size=0.8, shape=19)+geom_hex(bins=50)+
 scale_fill_viridis(option="plasma") + theme_bw()+
   facet_wrap( ~ treatment)+
  geom_smooth(method = lm)+
  xlab("ESRP2 HCR Spot Number per Cell") +
  ylab("ESRP1 HCR Spots Number per Cell")+
  theme(
            strip.text.x = element_text(size = 14),
            legend.title = element_text(size=14), #change legend title font size
            legend.text = element_text(size=12, face="bold"), #change legend text font size
            axis.title = element_text(size = 14, color="black", face="bold"),
            axis.text = element_text(size=12,color="black", face="bold"))
  
```


Document the information about the analysis session
```{r sessionInfo, include=TRUE, echo=TRUE, results='markup'}
sessionInfo()
