---
title: "Figures S2A, S2B, S2C, and S2D"
author: "Asaf Shilo/Gianluca Pegoraro"
date: "Aug 31st 2022"
output: github_document
---

### Read and Process the Object Level Data

Load required packages.
```{r loadAndClean}
library(plyr)
library(tidyverse)
library(data.table)
library(stringr)
library(ggthemes)
library(ineq)
library(ggridges)
library(viridis)
library(readr)
library(curl)
```

Set Uotput folder

```{r, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  fig.path = 'output/',
  message = FALSE,
  warning = FALSE
  )
```

Set the palette and the running theme for ggplot2.

```{r setThemeandPalette}
theme_set(theme_minimal())
theme_update(axis.text.x = element_text(
angle = -90,
hjust = 0,
vjust = 0.5
))
```

Read and Change variable types in experimental metadata data.table 
```{r readMetadata}
dt_md <- fread("metadata/Layout.txt")
dt_md <- dt_md[probe != "",]
dt_md[, `:=`(probe = factor(probe, levels = c("PB1",
                                              "PB2",
                                              "PB3",
                                              "PB4",
                                              "PB1 PB4",
                                              "PB2 PB3",
                                              "PB1 PB3 PB4",
                                              "PB2 PB3 PB4",
                                              "PB All",
                                              "PC1",
                                              "PC2",
                                              "PC3",
                                              "PC4",
                                              "PC1 PC4",
                                              "PC2 PC3",
                                              "PC1 PC3 PC4",
                                              "PC2 PC3 PC4",
                                              "PC All",
                                              "None"
                                              )))]
```

### Download the data if needed

Download unzip the Columbus results of the experiments from Figshare if they have not been already downloaded.

```{r downlaod-data}
if(!dir.exists("ObjectLevelData")) {
  URL <- "https://figshare.com/ndownloader/files/##35015029##" 
  curl_download(URL, "input.zip")
  unzip("input.zip")
}
```


### Read and process the cell level data

Set RegEx patterns for directory searches for cell level data.

```{r regexFilename}
pat_c <- ".*Nuclei Final\\[0\\]\\.txt$" # Pattern for Single Cell data files
```

Create a list of the RegEx patterns set in the previous chunk. **Important:** the list names will be carried over all the next steps!!!

-   c = cell level data

```{r patList}
pat_list <- list(c= pat_c)
```

Recursively search the `ObjectLevelData` directory and its subdirectories for files whose name includes the RegEx patterns defined two chunks above. The `path_list` functon outputs absolute file names. `path_list` is a list containing all the filenames on a per cell-level.
```{r directorySearch}
list_files <- function(x) {
  dir(
  path = 'ObjectLevelData/',
  pattern = x,
  full.names = TRUE,
  recursive = TRUE,
  include.dirs = TRUE
  )
}

path_list <- llply(pat_list, list_files) 
```

Extract file names from absolut path and set them as list element names. 

```{r trimNames}
trim_names <- function(x) {
  names(x) <-
  basename(x) # This assigns the filename to the file that it is read
  y <- x ## This is necessary because of scoping issues
}

path_list <- llply(path_list, trim_names) 
```

Recursively read and merge object level data files as data.frames. Rows are labeled with relative filenames (The `.id` variable). This and the previous chunks are slightly modified tricks adopted from H. Wickam ["Tidy Data" paper](http://vita.had.co.nz/papers/tidy-data.pdf).

```{r readMerge}
read_merge <- function(x) {
  dt <- as.data.table(ldply(x, fread, sep = "\t", na.strings = NULL))
}

dt_list <- llply(path_list, read_merge)
```

Separate the cell level data.

```{r deleteDtList}
dt_cell <- dt_list$c

rm(dt_list)
```

Change variables names.

```{r changeNames, skip_absent=TRUE}
setnames(
  dt_cell,
  c("Row",
    "Column",
    "ScreenName",
    "WellName",
    "Nuclei Final - Number of Positive 488- per Cell",
    "Nuclei Final - Intensity Spot 488 Mean- Mean per Cell",
    "Nuclei Final - Intensity Spot 488 Mean- Sum per Cell",
    "Nuclei Final - Spot Area [px²]- Mean per Cell",
    "Nuclei Final - Spot Area [px²]- Sum per Cell"
  ),
  c("row",
    "column",
    "plate",
    "well",
    "spot_488_n",
    "Mean_spot_intensity_488",
    "Sum_spot_intensity_488",
    "Mean_spot_area_488",
    "Sum_spot_area_488"
    )
 )
```

Join measurement data with experimental metadata.

```{r}
setkey(dt_md, row, column)
setkey(dt_cell, row, column)

dt_full <- dt_cell[dt_md, nomatch = 0] 
```

Generate datasets for plotting.

```{r}
dt_T47D <- filter(dt_full,cell_line == "T47D" & probe !="None")
dt_T47D_PB <- dt_T47D%>%filter(!grepl('PC', probe))
dt_T47D_PC <- dt_T47D%>%filter(!grepl('PB', probe))                              
dt_AN3 <- filter(dt_full,cell_line == "AN3" & probe !="None")
dt_AN3_PC <- dt_AN3%>%filter(!grepl('PB', probe))
dt_AN3_PB <- dt_AN3%>%filter(!grepl('PC', probe))
```

### Plotting

Plotting FGFR2_IIIB sum spot area in T47D cells

```{r FigS2_A, echo = F}
sarea_spot_T47D_PB <- ggplot(dt_T47D_PB, aes(x = probe, y = Sum_spot_area_488, fill = probe_n))

sarea_spot_T47D_PB + geom_violin(alpha=1, width=0.75) + geom_boxplot(outlier.shape = NA, width=0.1, color="black", fill = "white", alpha=0.8)+
            
            scale_fill_distiller(guide = guide_legend(title = "Number of\n probes", title.hjust = 0.5 ))+
            ylim(0,750)+
  
            labs(title = "FGFR2-IIIb Expression in T-47D cells", x = "probe set", y = "Sum FGFR2-IIIb HCR Spots area\n per cell") +
  
            theme_minimal()+
                  theme(
                  plot.title = element_text(size = 16, hjust = 0.5),
                  axis.text.x = element_text(angle = 45, hjust = 1),
                  axis.title = element_text(size = 12, color = "black", face = "bold"),
                  axis.text = element_text(size=10,color="black", face="bold"),
                  legend.title = element_text(colour="black", size=8, face="bold")
                  )
```

Plotting FGFR2_IIIB mean spot area in T47D cells

```{r FigS2_B, echo = F}
marea_spot_T47D_PB <- ggplot(dt_T47D_PB, aes(x = probe, y = Mean_spot_area_488, fill = probe_n))

marea_spot_T47D_PB + geom_violin(alpha=1, width=0.75) + geom_boxplot(outlier.shape = NA, width=0.1, color="black", fill = "white", alpha=0.8)+
            
            scale_fill_distiller(guide = guide_legend(title = "Number of\n probes", title.hjust = 0.5 ))+
            ylim(10,35)+
            labs(title = "FGFR2-IIIb Expression in T-47D cells", x = "probe set", y = "Mean FGFR2-IIIb HCR Spots area\n per cell (px^2)") +
            theme_minimal()+
                  theme(
                  plot.title = element_text(size = 16, hjust = 0.5),
                  axis.text.x = element_text(angle = 45, hjust = 1),
                  axis.title = element_text(size = 12, color = "black", face = "bold"),
                  axis.text = element_text(size=10,color="black", face="bold"),
                  legend.title = element_text(colour="black", size=8, face="bold")
                  )
```

Plotting FGFR2_IIIc sum spot area in AN3CA cells

```{r FigS2_C, echo = F}
sarea_spot_AN3_PB <- ggplot(dt_AN3_PC, aes(x = probe, y = Sum_spot_area_488, fill = probe_n))

sarea_spot_AN3_PB + geom_violin(alpha=1, width=0.75) + geom_boxplot(outlier.shape = NA, width=0.1, color="black", fill = "white", alpha=0.8)+
            
            scale_fill_distiller(guide = guide_legend(title = "Number of\n probes", title.hjust = 0.5 ))+
            ylim(0,1500)+
  
            labs(title = "FGFR2-IIIc Expression in AN3 CA cells", x = "probe set", y = "Sum FGFR2-IIIc HCR Spots area\n per cell") +
  
            theme_minimal()+
                  theme(
                  plot.title = element_text(size = 16, hjust = 0.5),
                  axis.text.x = element_text(angle = 45, hjust = 1),
                  axis.title = element_text(size = 12, color = "black", face = "bold"),
                  axis.text = element_text(size=10,color="black", face="bold"),
                  legend.title = element_text(colour="black", size=8, face="bold")
                  )
```

Plotting FGFR2_IIIc mean spot area in AN3CA cells

```{r FigS2_D, echo = F}
marea_spot_AN3_PC <- ggplot(dt_AN3_PC, aes(x = probe, y = Mean_spot_area_488, fill = probe_n))

marea_spot_AN3_PC + geom_violin(alpha=1, width=0.75) + geom_boxplot(outlier.shape = NA, width=0.1, color="black", fill = "white", alpha=0.8)+
            
            scale_fill_distiller(guide = guide_legend(title = "Number of\n probes", title.hjust = 0.5 ))+
            ylim(10,35)+
  
            labs(title = "FGFR2-IIIc Expression in AN3 CA cells", x = "probe set", y = "Mean FGFR2-IIIc HCR Spots area\n per cell (px^2)") +
  
            theme_minimal()+
                  theme(
                  plot.title = element_text(size = 16, hjust = 0.5),
                  axis.text.x = element_text(angle = 45, hjust = 1),
                  axis.title = element_text(size = 12, color = "black", face = "bold"),
                  axis.text = element_text(size=10,color="black", face="bold"),
                  legend.title = element_text(colour="black", size=8, face="bold")
                  )
```

Document the information about the analysis session

```{r sessionInfo, include=TRUE, echo=TRUE, results='markup'}
sessionInfo()
