---
title: "Figure S2A"
author: "Asaf Shilo/Gianluca Pegoraro"
date: "May 12th 2022"
output: github_document
---

### Analysis setup

Load required packages.

```{r loadAndClean}
library(plyr)
library(tidyverse)
library(data.table)
library(stringr)
library(ggthemes)
library(dplyr)
library(curl)
```

```{r, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  fig.path = 'output/',
  message = FALSE,
  warning = FALSE
  )
```

Set the palette and the running theme for ggplot2.

```{r setThemeandPalette}
theme_set(theme_minimal())
theme_update(axis.text.x = element_text(
angle = -90,
hjust = 0,
vjust = 0.5
))
```

### Read the experimental metadata

Read and change variable types in experimental metadata data.table.

```{r readMetadata}
dt_md <- fread("metadata/layout.txt")
dt_md <- dt_md[probe != "",]
dt_md[, `:=`(probe = factor(probe, levels = c("IIIB_B1",
                                              "IIIB_B3",
                                              "None")))]
```

### Download the data if needed

Download unzip the Columbus results of the experiments from Figshare if they have not been already downloaded.

```{r downlaod-data}
if(!dir.exists("ObjectLevelData")) {
  URL <- ""
  curl_download(URL, "input.zip")
  unzip("input.zip")
}
```

### Read and process the cell level data

Set RegEx patterns for directory searches for cell level data.

```{r regexFilename}
pat_c <- ".*Nuclei Final\\[0\\]\\.txt$" # Pattern for Single Cell data files
```

Create a list of the RegEx patterns set in the previous chunk. **Important:** the list names will be carried over all the next steps!!!

-   c = cell level data

```{r patList}
pat_list <- list(c= pat_c)
```

Recursively search the `ObjectLevelData` directory and its sub-directories for files whose name includes the RegEx patterns defined two chunks above. The `path_list` function outputs absolute file names. `path_list` is a list containing all the file names on a per cell-level.

```{r directorySearch}
list_files <- function(x) {
  dir(
  path = 'ObjectLevelData/',
  pattern = x,
  full.names = TRUE,
  recursive = TRUE,
  include.dirs = TRUE
  )
}

path_list <- llply(pat_list, list_files) 
```

Extract file names from absolute path and set them as list element names.

```{r trimNames}
trim_names <- function(x) {
  names(x) <-
  basename(x) # This assigns the filename to the file that it is read
  y <- x ## This is necessary because of scoping issues
}

path_list <- llply(path_list, trim_names) 
```

Recursively read and merge object level data files as data.frames. Rows are labeled with relative filenames (The `.id` variable). This and the previous chunks are slightly modified tricks adopted from H. Wickam ["Tidy Data" paper](http://vita.had.co.nz/papers/tidy-data.pdf).

```{r readMerge}
read_merge <- function(x) {
  dt <- as.data.table(ldply(x, fread, sep = "\t", na.strings = NULL))
}

dt_list <- llply(path_list, read_merge)
```

Separate the cell level data.

```{r deleteDtList}
dt_cell <- dt_list$c

rm(dt_list)
```

Change variables names.

```{r changeNames}
setnames(
  dt_cell,
  c("Row",
    "Column",
    "ScreenName",
    "WellName",
    "Nuclei Final - Number of Positive 647- per Cell",
    "Nuclei Final - Number of Positive 488- per Cell"
  ),
  c("row",
    "column",
    "plate",
    "well",
    "B3_Far_Red",
    "B1_Green")
  )
```

Join measurement data with experimental metadata.

```{r}
setkey(dt_md, row, column)
setkey(dt_cell, row, column)

dt_full <- dt_cell[dt_md, nomatch = 0] 
```

Generate datasets for plotting.

```{r subset}
dt_ctrl <- subset(dt_full, Treatment != "si_FGFR2")
dt_1 <- subset(dt_ctrl, probe == "IIIB_B1")
dt_2 <- dt_1
names (dt_2) [22] <-'spot_N'
dt_3 <-select (dt_2, c(8:13, 22:25))

dt_1.1 <- subset(dt_ctrl, probe == "IIIB_B3")
dt_2.1 <- dt_1.1
names (dt_2.1) [21] <-'spot_N'
dt_3.1 <-select (dt_2.1, c(8:13,21, 23:25))

dt_all <- rbind(dt_3, dt_3.1)
```

### Plotting

Plotting FGFRII-IIIB data from T-47D cells.

```{r FigS2_A, echo = F}
spot_n_boxplot <- ggplot(dt_all, aes(x = probe, y = spot_N))

spot_n_boxplot + geom_violin(fill = "green4") + geom_boxplot(width=0.075, color="black", fill ="white", outlier.shape = NA)+
  ylim(0,40)+
  scale_y_continuous(expand = c(0, 0), limits = c(0, 30))+
   labs(title = "FGFR2 IIIb Expression in T-47D cells", y = "HCR Spots per cell") +
  scale_fill_manual(values = c("green", "orange"))+
  theme_classic()+
                        theme(
                legend.position ="none",
                plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
                axis.title.x = element_blank(),
                axis.title = element_text(size = 14, color="black", face="bold"),
                axis.text = element_text(size=14,color="black", face="bold"))
```

Document the information about the analysis session. 

```{r sessionInfo, include=TRUE, echo=TRUE, results="markup"}
sessionInfo()
```
