---
title: "Fig. 4 no"
author: "Asaf Shilo"
date: "March 16th 2022"
output: html_document
---

### Read and Process the Object Level Data

Load required packages.
```{r loadAndClean}
library(plyr)
library(tidyverse)
library(data.table)
library(stringr)
library(ggthemes)
library(ggpubr)
```

```{r, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  fig.path = 'R_Output_Fig4/',
  dev = c("png", "tiff", "pdf"),
  message = FALSE,
  warning = FALSE
  )
```


Read and Change variable types in experimental metadata data.table
```{r readMetadata}
dt_md <- fread("ExperimentalMetaData/GlobalLayoutplusMCF.txt")
dt_md <- dt_md[probe != "",]
dt_md[, `:=`(probe = factor(probe, levels = c("None", 
                                              "FGFR2",
                                              "FGFR2-IIIb",
                                              "PGK1",
                                              "FGFR2-D7-10",
                                              "All" )))]
```

Set RegEx patterns for directory searches for cell level data.
```{r regexFilename}
pat_c <- ".*Nuclei Final\\[0\\]\\.txt$" # Pattern for Single Cell data files
```

Create a list of the RegEx patterns set in the previous chunk. **Important:** the list names will be carried over all the next steps!!!

* c = cell level data

```{r patList}
pat_list <- list(c= pat_c)
```

Recursively search the `ObjectLevelData` directory and its subdirectories for files whose name includes the RegEx patterns defined two chunks above. The `path_list` functon outputs absolute file names. `path_list` is a list containing all the filenames on a per cell-level.
```{r directorySearch}
list_files <- function(x) {
  dir(
  path = 'ObjectLevelData/',
  pattern = x,
  full.names = TRUE,
  recursive = TRUE,
  include.dirs = TRUE
  )
}

path_list <- llply(pat_list, list_files) 
```

Extract file names from absolut path and set them as list element names. 
```{r trimNames}
trim_names <- function(x) {
  names(x) <-
  basename(x) # This assigns the filename to the file that it is read
  y <- x ## This is necessary because of scoping issues
}

path_list <- llply(path_list, trim_names) 
```

Recursively read and merge object level data files as data.frames. Rows are labeled with relative filenames (The `.id` variable). This and the previous chunks are slightly modified tricks adopted from H. Wickam ["Tidy Data" paper](http://vita.had.co.nz/papers/tidy-data.pdf).
```{r readMerge}
read_merge <- function(x) {
  dt <- as.data.table(ldply(x, fread, sep = "\t", na.strings = NULL))
}

dt_list <- llply(path_list, read_merge)
```

Separate the cell level data from the distance and spot level data.
```{r deleteDtList}
dt_cell <- dt_list$c

rm(dt_list)
```

Change attribute names and create a new `im_acq` variable to indicate 2D or 3D imaging acquisitions.
```{r changeNames}
setnames(
  dt_cell,
  c("Row",
    "Column",
    "ScreenName",
    "WellName",
    
    "Nuclei Final - Nucleus Roundness",
    "Nuclei Final - Number of Positive 488- per Cell",
    "Nuclei Final - Number of Positive 562- per Cell",
    "Nuclei Final - Number of Positive 647- per Cell"),
  c("row",
    "column",
    "plate",
    "well",
   
    "roundness",
    "spot_n_488",
    "spot_n_562",
    "spot_n_647")
  )
```

Join measurementa data with experimental metadata.
```{r}
setkey(dt_md, row, column)
setkey(dt_cell, row, column)

dt_full <- dt_cell[dt_md, nomatch = 0] 
```

create df
```{r}
dt_MP<- filter(dt_full, probe !='FGFR2')
```


plots for ALL WITH 488 CHANNEL
```{r, fig.width=10,fig.height=9}
spot_488_boxplot <- ggplot(dt_MP, aes(x = probe, y = spot_n_488, fill = cell_line))

spot_488_boxplot + geom_boxplot() +
          ylim(0,25)+
          labs(title = "FGFR2-IIIb", y = "FGFR2-IIIb HCR Spots per cell") +
                theme_classic()+
                theme(
                legend.position =c(0.15, 0.9),
                legend.title = element_blank(),
                legend.text = element_text(size=18),
                plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
                axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 20, color="black", face="bold"),
                axis.text = element_text(size=20,color="black", face="bold"))
```


plots for ALL WITH 562 CHANNEL
```{r, fig.width=10,fig.height=9}
spot_562_boxplot <- ggplot(dt_MP, aes(x = probe, y = spot_n_562, fill = cell_line))

spot_562_boxplot + geom_boxplot() +
          ylim(0,200)+
          labs(title = "PGK1" ,y = "PGK1 HCR Spots per cell") +
                theme_classic()+
                theme(
                legend.position =c(0.15, 0.9),
                legend.title = element_blank(),
                legend.text = element_text(size=18),
                plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
                axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 20, color="black", face="bold"),
                axis.text = element_text(size=20,color="black", face="bold"))
```


plots for ALL WITH 647 CHANNEL
```{r, fig.width=10,fig.height=9}
spot_647_boxplot <- ggplot(dt_MP, aes(x = probe, y = spot_n_647, fill = cell_line))

spot_647_boxplot + geom_boxplot() +
          ylim(0,100)+
          labs(title  = "FGFR2-D7-10", y = "FGFR2-D7-10 HCR Spots per cell") +
                theme_classic()+
                theme(
                legend.position =c(0.15, 0.9),
                legend.title = element_blank(),
                legend.text = element_text(size=18),
                plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
                axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 20, color="black", face="bold"),
                axis.text = element_text(size=20,color="black", face="bold"))
```


create df with IIIb OE only
```{r}
dt_MP_IIIbOE <- filter(dt_MP, probe == 'All' & cell_line == 'U2OS FGFR2-IIIb')
```


```{r, }
spot_FGFR2_scatter <- ggplot(dt_MP_IIIbOE, aes(x = spot_n_647, y = spot_n_488))

spot_FGFR2_scatter + geom_point() +
          geom_smooth(method = lm)+
          
          stat_regline_equation(label.x=04, label.y=195, size=5, face = "bold") +
          stat_cor(label.x=04, label.y=175, size = 5, face = "bold")+
            xlim(0,150)+
            ylim(0,200)+
            labs(x = "FGFR2-D7-10 HCR Spots per Cell ", y = "FGFR2-IIIb HCR Spots per Cell") +
            theme_classic()+
                theme(
                plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 14, color="black", face="bold"),
                axis.text = element_text(size=12,color="black", face="bold"))
```

```{r, }
spot_FGFR2_scatter <- ggplot(dt_MP_IIIbOE, aes(x = spot_n_647, y = spot_n_488))

spot_FGFR2_scatter + geom_point() +
          geom_smooth(method = lm)+
          
         
            xlim(0,150)+
            ylim(0,200)+
            labs(x = "FGFR2-D7-10 HCR Spots per Cell ", y = "FGFR2-IIIb HCR Spots per Cell") +
            theme_classic()+
                theme(
                plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 14, color="black", face="bold"),
                axis.text = element_text(size=12,color="black", face="bold"))
```

Document the information about the analysis session
```{r sessionInfo, include=TRUE, echo=TRUE, results='markup'}
sessionInfo()