---
title: "Figures S1A, S1B, S1C, and S1D"
author: "Asaf Shilo/Gianluca Pegoraro"
date: "May 12th 2022"
output: github_document
---

### Analysis setup

Load required packages.

```{r loadAndClean}
library(plyr)
library(tidyverse)
library(data.table)
library(stringr)
library(ggthemes)
library(ineq)
library(ggridges)
library(scales)
library(curl)
```

Set output folder.

```{r, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  fig.path = 'output/',
  message = FALSE,
  warning = FALSE
  )
```

Set the palette and the running theme for ggplot2.

```{r setThemeandPalette}
theme_set(theme_minimal())
theme_update(axis.text.x = element_text(
angle = -90,
hjust = 0,
vjust = 0.5
))
```

### Read the experimental metadata

Read and change variable types in experimental metadata data.table.

```{r readMetadata}
dt_md <- fread("metadata/layout.txt")
dt_md <- dt_md[probe != "",]
dt_md[, `:=`(probe = factor(probe, levels = c("FGFR2-IIIb", 
                                              "FGFR2-IIIc",
                                              "FGFR2-Full",
                                              "All",
                                              "None")))]
```

### Download the data if needed

Download unzip the Columbus results of the experiments from Figshare if they have not been already downloaded.

```{r downlaod-data}
if(!dir.exists("ObjectLevelData")) {
  URL <- ""
  curl_download(URL, "input.zip")
  unzip("input.zip")
}
```

### Read and process the cell level data

Set RegEx patterns for directory searches for cell level data.

```{r regexFilename}
pat_c <- ".*Nuclei Final\\[0\\]\\.txt$" # Pattern for Single Cell data files
```

Create a list of the RegEx patterns set in the previous chunk. **Important:** the list names will be carried over all the next steps!!!

-   c = cell level data

```{r patList}
pat_list <- list(c= pat_c)
```

Recursively search the `ObjectLevelData` directory and its subdirectories for files whose name includes the RegEx patterns defined two chunks above. The `path_list` functon outputs absolute file names. `path_list` is a list containing all the filenames on a per cell-level.

```{r directorySearch}
list_files <- function(x) {
  dir(
  path = 'ObjectLevelData/',
  pattern = x,
  full.names = TRUE,
  recursive = TRUE,
  include.dirs = TRUE
  )
}

path_list <- llply(pat_list, list_files) 
```

Extract file names from absolute path and set them as list element names.

```{r trimNames}
trim_names <- function(x) {
  names(x) <-
  basename(x) # This assigns the filename to the file that it is read
  y <- x ## This is necessary because of scoping issues
}

path_list <- llply(path_list, trim_names) 
```

Recursively read and merge object level data files as data.frames. Rows are labeled with relative filenames (The `.id` variable). This and the previous chunks are slightly modified tricks adopted from H. Wickam ["Tidy Data" paper](http://vita.had.co.nz/papers/tidy-data.pdf).

```{r readMerge}
read_merge <- function(x) {
  dt <- as.data.table(ldply(x, fread, sep = "\t", na.strings = NULL))
}

dt_list <- llply(path_list, read_merge)
```

Separate the cell level data.

```{r deleteDtList}
dt_cell <- dt_list$c

rm(dt_list)
```

Change variables names.

```{r changeNames}
setnames(
  dt_cell,
  c("Row",
    "Column",
    "ScreenName",
    "WellName",
    "Nuclei Final - Number of Positive 647- per Cell",
    "Nuclei Final - Number of Positive 562- per Cell",
    "Nuclei Final - Number of Positive 488- per Cell",
    "Nuclei Final - Nucleus Roundness",
    "Nuclei Final - Intensity Spot 488 Mean- Sum per Cell",
    "Nuclei Final - Intensity Spot 561 Mean- Sum per Cell",
    "Nuclei Final - Intensity Spot 640 Mean- Sum per Cell"
  ),
  c("row",
    "column",
    "plate",
    "well",
    "spot_647_n",
    "spot_562_n",
    "spot_488_n",
    "roundness",
    "Intensity_488",
    "Intensity_562",
    "Intensity_647"
    )
  )
```

Join measurement data with experimental metadata.

```{r}
setkey(dt_md, row, column)
setkey(dt_cell, row, column)

dt_full <- dt_cell[dt_md, nomatch = 0] 
```

Generate datasets for plotting.

```{r, echo = F}
All_dt <-filter (dt_full, probe == "All")
Full_dt<- filter(dt_full,probe=="FGFR2-Full")
IIIB_dt <- filter (dt_full,probe=="FGFR2-IIIb")
IIIC_dt <- filter (dt_full,probe=="FGFR2-IIIc")
```

### Plotting

FGFR2-Full probes - Box plot for spot count

```{r FigS1_A_left, fig.width=8, fig.height=8, echo=FALSE}
spot_Full_box <- ggplot(Full_dt, aes(y = spot_647_n, x = cell_line))

spot_Full_box +
  geom_boxplot(alpha = 0.8,
               color = "black",
               fill = "red3") +
  ylim(0, 100) +
  scale_color_tableau(name = "Cell Line") +
  scale_fill_tableau(name = "Cell Line") +
  labs(title = "FGFR2-Full",
       x = "x title",
       y = "FGFR2-Full HCR Spot Number per Cell") +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.text = element_text(size = 16),
    legend.position = "none",
    plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(size = 20, color = "black", face =
                                "bold"),
    axis.text = element_text(size = 20, color = "black", face =
                               "bold")
  )
```

FGFR2-Full probes - Box plot for intensity

```{r FigS1_A_right, fig.width=8, fig.height=8, echo=FALSE}
Intensity_Full_boxplot <-
  ggplot(Full_dt, aes(y = Intensity_647 , x = cell_line))

Intensity_Full_boxplot +
  geom_boxplot(alpha = 0.8,
               color = "black",
               fill = "red3") +
  scale_y_continuous(labels = scientific) +
  ylim(0, 30000) +
  scale_color_tableau(name = "Cell Line") +
  scale_fill_tableau(name = "Cell Line") +
  labs(title = "FGFR2-Full",
       x = "x title",
       y = "Mean HCR Spot Intensity - Sum per Cell (A.U.)") +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.text = element_text(size = 16),
    legend.position = "none",
    plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(size = 20, color = "black", face =
                                "bold"),
    axis.text = element_text(size = 20, color = "black", face =
                               "bold")
  )
```

FGFR2-IIIb probes - Box plot for spot count

```{r FigS1_B_left, fig.width=8, fig.height=8, echo=FALSE}
spot_IIIb_box <-
  ggplot(IIIB_dt, aes(y = spot_488_n , x = cell_line))

spot_IIIb_box +
  geom_boxplot(alpha = 0.8,
               color = "black",
               fill = "green4") +
  ylim(0, 20) +
  scale_color_tableau(name = "Cell Line") +
  scale_fill_tableau(name = "Cell Line") +
  labs(title = "FGFR2-IIIb",
       x = "x title",
       y = "FGFR2-IIIb HCR Spot Number per Cell") +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.text = element_text(size = 16),
    legend.position = "none",
    plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(size = 20, color = "black", face =
                                "bold"),
    axis.text = element_text(size = 20, color = "black", face =
                               "bold")
  )
```

FGFR2-IIIb probes - Box plot for intensity

```{r FigS1_B_right, fig.width=8, fig.height=8, echo=FALSE}
Intensity_IIIb_boxplot <- ggplot(IIIB_dt, aes(y = Intensity_488 , x= cell_line)) 
Intensity_IIIb_boxplot + 
  geom_boxplot(alpha = 0.8, color = "black", fill = "green4") + 
  scale_y_continuous(labels = scientific)+
   ylim(0,15000)+
                scale_color_tableau(name = "Cell Line") +
                scale_fill_tableau(name = "Cell Line") +
                
                labs(title="FGFR2-IIIb",
                      x ="x title", 
                      y = "Mean HCR Spot Intensity - Sum per Cell (A.U.)")+
  
                
                theme_classic()+
                theme(
                legend.title = element_blank(),
                legend.text = element_text(size=16),
                legend.position ="none",
                plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
                axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 20, color="black", face="bold"),
                axis.text = element_text(size=20,color="black", face="bold"))
  
```

FGFR2-IIIc probes - Box plot for spot count

```{r FigS1_C_left, fig.width=8, fig.height=8, echo=FALSE}
spot_IIIc_box <-
  ggplot(IIIC_dt, aes(y = spot_488_n , x = cell_line))

spot_IIIc_box + geom_boxplot(alpha = 0.8,
                             color = "black",
                             fill = "blue3") +
  
  ylim(0, 35) +
  scale_color_tableau(name = "Cell Line") +
  scale_fill_tableau(name = "Cell Line") +
  labs(title = "FGFR2-IIIc",
       x = "x title",
       y = "FGFR2-IIIb HCR Spot Number per Cell") +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.text = element_text(size = 16),
    legend.position = "none",
    plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(size = 20, color = "black", face =
                                "bold"),
    axis.text = element_text(size = 20, color = "black", face =
                               "bold")
  )
```

FGFR2-IIIc probes - Box plot for intensity

```{r FigS1_C_right, fig.width=8, fig.height=8, echo=FALSE}
Intensity_IIIc_boxplot <-
  ggplot(IIIC_dt, aes(y = Intensity_488 , x = cell_line))
Intensity_IIIc_boxplot + geom_boxplot(alpha = 0.8,
                                      color = "black",
                                      fill = "blue3") +
  scale_y_continuous(labels = scientific) +
  ylim(0, 25000) +
  scale_color_tableau(name = "Cell Line") +
  scale_fill_tableau(name = "Cell Line") +
  
  labs(title = "FGFR2-IIIc",
       x = "x title",
       y = "Mean HCR Spot Intensity - Sum per Cell (A.U.)") +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.text = element_text(size = 16),
    legend.position = "none",
    plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(size = 20, color = "black", face =
                                "bold"),
    axis.text = element_text(size = 20, color = "black", face =
                               "bold")
  )
```

TBP probes - Box plot for spot count

```{r FigS1_D_left, fig.width=8, fig.height=8, echo=FALSE}
spot_TBP_box <- ggplot(All_dt, aes(y = spot_562_n , x = cell_line))

spot_TBP_box + geom_boxplot(alpha = 0.8,
                            color = "black",
                            fill = "#00BFC4") +
  ylim(0, 50) +
  scale_color_tableau(name = "Cell Line") +
  scale_fill_tableau(name = "Cell Line") +
  
  labs(title = "TBP",
       x = "x title",
       y = "TBP HCR Spot Number per Cell") +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.text = element_text(size = 16),
    legend.position = "none",
    plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(size = 20, color = "black", face =
                                "bold"),
    axis.text = element_text(size = 20, color = "black", face =
                               "bold")
  )
```

TBP probes - Box plot for intensity

```{r FigS1_D_right, fig.width=8, fig.height=8, echo=FALSE}
Intensity_TBP_boxplot <-
  ggplot(All_dt, aes(y = Intensity_562 , x = cell_line))

Intensity_TBP_boxplot + geom_boxplot(alpha = 0.8,
                                     color = "black",
                                     fill = "#00BFC4") +
  scale_y_continuous(labels = scientific) +
  ylim(0, 10000) +
  scale_color_tableau(name = "Cell Line") +
  scale_fill_tableau(name = "Cell Line") +
  
  labs(title = "TBP",
       x = "x title",
       y = "Mean HCR Spot Intensity - Sum per Cell (A.U.)") +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.text = element_text(size = 16),
    legend.position = "none",
    plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(size = 20, color = "black", face =
                                "bold"),
    axis.text = element_text(size = 20, color = "black", face =
                               "bold")
  )
```

Document the information about the analysis session. 

```{r sessionInfo, include=TRUE, echo=TRUE, results="markup"}
sessionInfo()
```
