---
title: "Fig. 1 I-K"
author: "Asaf Shilo"
date: "March 16th 2022"
output: html_document
---

### Read and Process the Object Level Data

Load required packages.
```{r loadAndClean}
library(plyr)
library(tidyverse)
library(data.table)
library(stringr)
library(ggthemes)
library(ineq)
library(ggridges)
library(scales)
```

Set output folder
```{r, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  fig.path = 'R_Output_Fig1_Histogram/',
  dev = c("tiff", "pdf", "tiff"),
  message = FALSE,
  warning = FALSE
  )
```


Read and Change variable types in experimental metadata data.table
```{r readMetadata}
dt_md <- fread("ExperimentalMetaData/GlobalLayout.txt")
dt_md <- dt_md[probe != "",]
dt_md[, `:=`(probe = factor(probe, levels = c("FGFR2-IIIb", 
                                              "FGFR2-IIIc",
                                              "FGFR2-Full",
                                              "All",
                                              "None")))]
```

Set RegEx patterns for directory searches for cell level data.
```{r regexFilename}
pat_c <- ".*Nuclei Final\\[0\\]\\.txt$" # Pattern for Single Cell data files
```

Create a list of the RegEx patterns set in the previous chunk. **Important:** the list names will be carried over all the next steps!!!

* c = cell level data

```{r patList}
pat_list <- list(c= pat_c)
```

Recursively search the `ObjectLevelData` directory and its subdirectories for files whose name includes the RegEx patterns defined two chunks above. The `path_list` functon outputs absolute file names. `path_list` is a list containing all the filenames on a per cell-level.
```{r directorySearch}
list_files <- function(x) {
  dir(
  path = 'ObjectLevelData/',
  pattern = x,
  full.names = TRUE,
  recursive = TRUE,
  include.dirs = TRUE
  )
}

path_list <- llply(pat_list, list_files) 
```

Extract file names from absolut path and set them as list element names. 
```{r trimNames}
trim_names <- function(x) {
  names(x) <-
  basename(x) # This assigns the filename to the file that it is read
  y <- x ## This is necessary because of scoping issues
}

path_list <- llply(path_list, trim_names) 
```

Recursively read and merge object level data files as data.frames. Rows are labeled with relative filenames (The `.id` variable). This and the previous chunks are slightly modified tricks adopted from H. Wickam ["Tidy Data" paper](http://vita.had.co.nz/papers/tidy-data.pdf).
```{r readMerge}
read_merge <- function(x) {
  dt <- as.data.table(ldply(x, fread, sep = "\t", na.strings = NULL))
}

dt_list <- llply(path_list, read_merge)
```

Separate the cell level data 
```{r deleteDtList}
dt_cell <- dt_list$c

rm(dt_list)
```

Change attribute names 
```{r changeNames}
setnames(
  dt_cell,
  c("Row",
    "Column",
    "ScreenName",
    "WellName",
    "Nuclei Final - Number of Positive 647- per Cell",
    "Nuclei Final - Number of Positive 562- per Cell",
    "Nuclei Final - Number of Positive 488- per Cell",
    "Nuclei Final - Nucleus Roundness"
    
  ),
  c("row",
    "column",
    "plate",
    "well",
    "spot_647_n",
    "spot_562_n",
    "spot_488_n",
    "roundness"
    
    )
  )
```

Join measurementa data with experimental metadata.
```{r}
setkey(dt_md, row, column)
setkey(dt_cell, row, column)

dt_full <- dt_cell[dt_md, nomatch = 0] 
```


Creating probes datasets
```{r, echo = F}
FGFR2_Full_dt <- filter (dt_full, probe == "FGFR2-Full")
FGFR2_IIIB_dt <- filter (dt_full,probe=="FGFR2-IIIb")
FGFR2_IIIC_dt <- filter (dt_full,probe=="FGFR2-IIIc")
```


Histogram Plot For FGFR2_Full Only Probes 
```{r, fig1.Full, fig.width=8,fig.height=11.5}
spot_FGFR2_Full_histo <- ggplot(FGFR2_Full_dt, aes(x = spot_647_n)) 

spot_FGFR2_Full_histo + geom_histogram(aes(y=..density..), binwidth = 1, colour="black", fill ="#FF0000", boundary = 0)+
               geom_density( color="black", size = 1)+ 

                facet_grid(cell_line ~ .,scales = "free_y") +
                xlim(0,90) +
                
  
                labs(title="FGFR2-Full",
                      x ="FGFR2-Full HCR Spots per Cell", 
                      y = "Density")+
  
                theme_bw()+
                theme(
                plot.title = element_text(size = 10, hjust = 0.5, face = "bold"),
                strip.text = element_text(size = 8),
                legend.position ="none",
                axis.title = element_text(size = 8),
                axis.text.x = element_text(size=8,color="black"),
                axis.text.y = element_text(size=8,color="black"))
  
```


Histogram Plot For FGFR2_IIIb Only Probes 
```{r, fig1.IIIb, fig.width=8,fig.height=11.5}
spot_FGFR2_IIIb_histo <- ggplot(FGFR2_IIIB_dt, aes(x = spot_488_n)) 

spot_FGFR2_IIIb_histo + geom_histogram(aes(y=..density..), binwidth = 1, colour="black", fill ="#008300", boundary = 0)+
               geom_density(alpha=.2, color="black", size = 1)+ 

                facet_grid(cell_line ~ .,scales = "free_y") +
                xlim(0,45) +
                
  
                labs(title="FGFR2-IIIb",
                      x ="FGFR2-IIIb HCR Spots per Cell", 
                      y = "Density")+
  
                theme_bw()+
                theme(
                plot.title = element_text(size = 10, hjust = 0.5, face = "bold"),
                strip.text = element_text(size = 8),
                legend.position ="none",
                axis.title = element_text(size = 8),
                axis.text.x = element_text(size=8,color="black"),
                axis.text.y = element_text(size=8,color="black"))
  
```

Histogram Plot For FGFR2_IIIc Only Probes 
```{r, fig1.IIIc, fig.width=8,fig.height=11.5}
spot_FGFR2_IIIb_histo <- ggplot(FGFR2_IIIC_dt, aes(x = spot_488_n)) 

spot_FGFR2_IIIb_histo + geom_histogram(aes(y=..density..), binwidth = 1, colour="black", fill ="#0000C8", boundary = 0)+
               geom_density(alpha=.2, color="black", size = 1)+ 

                facet_grid(cell_line ~ .,scales = "free_y") +
                xlim(0,45) +
                
  
                labs(title="FGFR2-IIIc",
                      x ="FGFR2-IIIc HCR Spots per Cell", 
                      y = "Density")+
  
                theme_bw()+
                theme(
                plot.title = element_text(size = 10, hjust = 0.5, face = "bold"),
                strip.text = element_text(size = 8),
                legend.position ="none",
                axis.title = element_text(size = 8),
                axis.text.x = element_text(size=8,color="black"),
                axis.text.y = element_text(size=8,color="black"))
  
```

Document the information about the analysis session
```{r sessionInfo, include=TRUE, echo=TRUE, results='markup'}
sessionInfo()
