---
title: "Figures S8B, S8C, S8D, and S8E"
author: "Asaf Shilo/Gianluca Pegoraro"
date: "August 31st 2022"
output: github_document
---

### Analysis setup

Load required packages.

```{r loadAndClean}
library(plyr)
library(tidyverse)
library(data.table)
library(stringr)
library(ggthemes)
library(hexbin)
library(ggpubr)
library(ineq)
library(ggridges)
library(viridis)
library(curl)
```

Set output folder.

```{r, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  fig.path = 'output/',
  message = FALSE,
  warning = FALSE
  )
```

Set the palette and the running theme for ggplot2.

```{r setThemeandPalette}
theme_set(theme_minimal())
theme_update(axis.text.x = element_text(
angle = -90,
hjust = 0,
vjust = 0.5
))
```

Read and Change variable types in experimental metadata data.table

```{r readMetadata}
dt_md <- fread("metadata/layout.txt")

dt_md <- dt_md[probe != "",]
dt_md[, `:=`(probe = factor(probe, levels = c("None", 
                                              "FGFR2-5'",
                                              "FGFR2-3'",
                                               "Both"
                                              )))]
```

### Download the data if needed

Download unzip the Columbus results of the experiments from Figshare if they have not been already downloaded.

```{r downlaod-data}
if(!dir.exists("ObjectLevelData")) {
  URL <- "https://figshare.com/ndownloader/files/35015029"
  curl_download(URL, "input.zip")
  unzip("input.zip")
}
```

### Read and process the cell level data

Set RegEx patterns for directory searches for cell level data.

```{r regexFilename}
pat_c <- ".*Nuclei Final\\[0\\]\\.txt$" # Pattern for Single Cell data files
```

Create a list of the RegEx patterns set in the previous chunk. **Important:** the list names will be carried over all the next steps!!!

-   c = cell level data

```{r patList}
pat_list <- list(c= pat_c)
```

Recursively search the `ObjectLevelData` directory and its sub-directories for files whose name includes the RegEx patterns defined two chunks above. The `path_list` function outputs absolute file names. `path_list` is a list containing all the file names on a per cell-level.

```{r directorySearch}
list_files <- function(x) {
  dir(
  path = 'ObjectLevelData/',
  pattern = x,
  full.names = TRUE,
  recursive = TRUE,
  include.dirs = TRUE
  )
}

path_list <- llply(pat_list, list_files) 
```

Extract file names from absolute path and set them as list element names.

```{r trimNames}
trim_names <- function(x) {
  names(x) <-
  basename(x) # This assigns the filename to the file that it is read
  y <- x ## This is necessary because of scoping issues
}

path_list <- llply(path_list, trim_names) 
```

Recursively read and merge object level data files as data.frames. Rows are labeled with relative filenames (The `.id` variable). This and the previous chunks are slightly modified tricks adopted from H. Wickam ["Tidy Data" paper](http://vita.had.co.nz/papers/tidy-data.pdf).

```{r readMerge}
read_merge <- function(x) {
  dt <- as.data.table(ldply(x, fread, sep = "\t", na.strings = NULL))
}

dt_list <- llply(path_list, read_merge)
```

Separate the cell level data.

```{r deleteDtList}
dt_cell <- dt_list$c

rm(dt_list)
```

Change variables names.

```{r changeNames}
setnames(
  dt_cell,
  c("Row",
    "Column",
    "ScreenName",
    "WellName",
    "Nuclei Final - Number of Positive 647- per Cell",
    "Nuclei Final - Number of Positive 562- per Cell",
    "Nuclei Final - Number of Positive 488- per Cell",
    "Nuclei Final - Nucleus Roundness"
  ),
  c("row",
    "column",
    "plate",
    "well",
    "spot_647_n",
    "spot_562_n",
    "spot_488_n",
    "roundness"
    )
  )
```

Join measurement data with experimental metadata.

```{r}
setkey(dt_md, row, column)
setkey(dt_cell, row, column)

dt_full <- dt_cell[dt_md, nomatch = 0] 
```

Generate datasets for plotting.

```{r}
dt_Full <- filter(dt_full, probe!='full') 
```

### Plotting

Plot Spot number for the 568 channel. 

```{r FigS8_B, echo = F}
spot_562_boxplot <- ggplot(dt_Full, aes(x = probe, y = spot_562_n, fill = probe))

spot_562_boxplot + geom_boxplot() + 
          ylim(0,50)+
           ylab("HCR Spot Number in 562")
```

Plot Spot number for the 640 channel. 

```{r FigS8_C, echo = F}
spot_647_boxplot <- ggplot(dt_Full, aes(x = probe, y = spot_647_n, fill = probe))

spot_647_boxplot + geom_boxplot() + 
          ylim(0,50)+
           scale_color_tableau(name = "Cell Line") +
           xlab("probe") +
           ylab("HCR Spot Number in 647")
```

```{r}
dt_Full_Both <- filter(dt_Full, probe=='Both') 
```

Plot scatterplot for 640 vs. 568 number of spots only for samples labelled as "both".

```{r FigS8_D, echo = F}
spot_FGFR2_split_2 <- ggplot(dt_Full_Both, aes(x = spot_647_n, y =  spot_562_n))

spot_FGFR2_split_2 + geom_point(size=0.8, shape=19)+
  scale_fill_viridis(option = "plasma") + theme_bw()+geom_hex(bins=50)+
  geom_smooth(method = lm)+
  stat_cor(method = "spearman",label.x=04, label.y=175, size = 5, face = "bold")+
  theme(aspect.ratio=1)+
  xlab("FGFR2-3' HCR Spot Number per Cell") +
                 ylab("FGFR2-5' HCR Spots Number per Cell")
  
```

Plot scatterplot for 640 vs. 568 number of spots for all samples.

```{r FigS8_E, echo = F}
spot_FGFR2_split <- ggplot(dt_Full, aes(x = spot_647_n, y =  spot_562_n, color = probe))

spot_FGFR2_split + geom_point(size=0.8, shape=19)+
  facet_wrap(. ~ probe) +
  geom_smooth(method = lm)+
  theme(aspect.ratio=1)+
  theme(legend.position = "none")+
  xlab("FGFR2-3' HCR Spot Number per Cell") +
                 ylab("FGFR2-5' HCR Spots Number per Cell")
```

Document the information about the analysis session. 

```{r sessionInfo, include=TRUE, echo=TRUE, results="markup"}
sessionInfo()
```
