---
title: "Figures 3A, 3B, 3C, and 3E"
author: "Asaf Shilo/Gianluca Pegoraro"
date: "Aug 31st 2022"
output: github_document
---

### Analysis setup

Load required packages.

```{r loadAndClean}
library(plyr)
library(tidyverse)
library(data.table)
library(stringr)
library(ggthemes)
library(ggpubr)
library(curl)
library(rstatix)
library(FSA)
```

Set output folder.

```{r, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  fig.path = 'output/',
  message = FALSE,
  warning = FALSE
  )
```

Set the palette and the running theme for ggplot2.

```{r setThemeandPalette}
theme_set(theme_minimal())
theme_update(axis.text.x = element_text(
angle = -90,
hjust = 0,
vjust = 0.5))

theme_update(plot.title = element_text(hjust = 0.5))
```

### Read the experimental metadata

Read and change variable types in experimental metadata data.table.

```{r readMetadata}
dt_md <- fread("metadata/layout.txt")
dt_md <- dt_md[probe != "",]
dt_md[, `:=`(probe = factor(probe, levels = c("None", 
                                              "FGFR2",
                                              "FGFR2-IIIb",
                                              "PGK1",
                                              "FGFR2-D7-10",
                                              "All" )))]
```

### Download the data if needed

Download unzip the Columbus results of the experiments from Figshare if they have not been already downloaded.

```{r downlaod-data}
if(!dir.exists("ObjectLevelData")) {
  URL <- "https://figshare.com/ndownloader/files/35014492"
  curl_download(URL, "input.zip")
  unzip("input.zip")
}
```

### Read and process the cell level data

Set RegEx patterns for directory searches for cell level data.

```{r regexFilename}
pat_c <- ".*Nuclei Final\\[0\\]\\.txt$" # Pattern for Single Cell data files
```

Create a list of the RegEx patterns set in the previous chunk. **Important:** the list names will be carried over all the next steps!!!

-   c = cell level data

```{r patList}
pat_list <- list(c= pat_c)
```

Recursively search the `ObjectLevelData` directory and its sub-directories for files whose name includes the RegEx patterns defined two chunks above. The `path_list` function outputs absolute file names. `path_list` is a list containing all the file names on a per cell-level.

```{r directorySearch}
list_files <- function(x) {
  dir(
  path = 'ObjectLevelData/',
  pattern = x,
  full.names = TRUE,
  recursive = TRUE,
  include.dirs = TRUE
  )
}

path_list <- llply(pat_list, list_files) 
```

Extract file names from absolute path and set them as list element names.

```{r trimNames}
trim_names <- function(x) {
  names(x) <-
  basename(x) # This assigns the file name to the file that it is read
  y <- x ## This is necessary because of scoping issues
}

path_list <- llply(path_list, trim_names) 
```

Recursively read and merge object level data files as data.frames. Rows are labeled with relative filenames (The `.id` variable). This and the previous chunks are slightly modified tricks adopted from H. Wickam ["Tidy Data" paper](http://vita.had.co.nz/papers/tidy-data.pdf).

```{r readMerge}
read_merge <- function(x) {
  dt <- as.data.table(ldply(x, fread, sep = "\t", na.strings = NULL))
}

dt_list <- llply(path_list, read_merge)
```

Separate the cell level data.

```{r deleteDtList}
dt_cell <- dt_list$c

rm(dt_list)
```

Change variables names.

```{r changeNames}
setnames(
  dt_cell,
  c("Row",
    "Column",
    "ScreenName",
    "WellName",
    "Nuclei Final - Nucleus Roundness",
    "Nuclei Final - Number of Positive 488- per Cell",
    "Nuclei Final - Number of Positive 562- per Cell",
    "Nuclei Final - Number of Positive 647- per Cell"),
  c("row",
    "column",
    "plate",
    "well",
    "roundness",
    "spot_n_488",
    "spot_n_562",
    "spot_n_647")
  )
```

Join measurement data with experimental metadata.

```{r}
setkey(dt_md, row, column)
setkey(dt_cell, row, column)

dt_full <- dt_cell[dt_md, nomatch = 0] 
```

Generate datasets for plotting.

```{r}
dt_IIIB <- filter(dt_full, probe == 'FGFR2-IIIb')
dt_Full <- filter(dt_full, probe == 'FGFR2')
dt_PGK1 <- filter(dt_full, probe == 'PGK1')
dt_D7_10 <-filter(dt_full, probe == 'FGFR2-D7-10')
dt_All <- filter (dt_full, probe == 'All')
dt_MP<- filter(dt_full, probe !='FGFR2')
```

### Statistcal tests

Kruskal-Wallis and Dunn test for FGFR2-Full

```{r, fig.width=10,fig.height=8} 
Full.kruskal <- dt_Full %>% kruskal_test(spot_n_488 ~ cell_line)
Full.kruskal

Full.kruskal_s <- dt_Full %>% kruskal_effsize(spot_n_488 ~ cell_line)

# Pairwise comparisons using Dunn’s test:
Full_pwc <- dt_Full %>% 
  dunn_test(spot_n_488 ~ cell_line) 
Full_pwc


Full_pwc_p <- Full_pwc %>% add_xy_position(x = "cell_line")

```

Kruskal-Wallis and Dunn test for FGFR2-IIIb

```{r, fig.width=10,fig.height=8} 
IIIB.kruskal <- dt_IIIB %>% kruskal_test(spot_n_488 ~ cell_line)
IIIB.kruskal

IIIB.kruskal_s <- dt_IIIB %>% kruskal_effsize(spot_n_488 ~ cell_line)

# Pairwise comparisons using Dunn’s test:
IIIB_pwc <- dt_IIIB %>% 
  dunn_test(spot_n_488 ~ cell_line) 
IIIB_pwc


IIIB_pwc_p <- IIIB_pwc %>% add_xy_position(x = "cell_line")

```

Kruskal-Wallis and Dunn test for PGK1

```{r, fig.width=10,fig.height=8} 
PGK1.kruskal <- dt_PGK1 %>% kruskal_test(spot_n_562 ~ cell_line)
PGK1.kruskal

PGK1.kruskal_s <- dt_PGK1 %>% kruskal_effsize(spot_n_562 ~ cell_line)

# Pairwise comparisons using Dunn’s test:
PGK1_pwc <- dt_PGK1 %>% 
  dunn_test(spot_n_562 ~ cell_line) 
PGK1_pwc


PGK1_pwc_p <- PGK1_pwc %>% add_xy_position(x = "cell_line")

```

### Plotting

Plotting total FGFR2 in B1-FGFR2 System.

```{r Fig3_A,fig.width=10,fig.height=8, echo=FALSE}
spot_B1Full_violin <- ggplot(dt_Full, aes(x = cell_line, y = spot_n_488))


spot_B1Full_violin + geom_violin(fill = "red3") + geom_boxplot(outlier.shape = NA, width=0.03, color="black")+
            ylim(0,75)+
            labs(y = "FGFR2-Full HCR Spots per cell") +
                theme_classic()+
                theme(
                legend.position ="none",
                plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
                axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 20, color="black", face="bold"),
                axis.text = element_text(size=20,color="black", face="bold"))
```

Plotting total FGFR2 in B1-FGFR2 System with p-values

```{r Fig3_A_2,fig.width=10,fig.height=8, echo=FALSE} 

spot_B1Full_violin_pv <- ggplot(dt_Full, aes(x = cell_line, y = spot_n_488))


spot_B1Full_violin_pv + geom_violin(fill = "red3") + geom_boxplot(outlier.shape = NA, width=0.03, color="black")+
  stat_pvalue_manual(Full_pwc_p, hide.ns = TRUE) +
            labs(y = "FGFR2-Full HCR Spots per cell")+
                theme_classic()+
                theme(
                legend.position ="none",
                plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
                axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 10, color="black", face="bold"),
                axis.text = element_text(size=10,color="black", face="bold"))
```

Plotting FGFR2-IIIb in B1-IIIB System

```{r Fig3_B,fig.width=10,fig.height=8, echo=FALSE}
spot_B1IIIb_violin <- ggplot(dt_IIIB, aes(x = cell_line, y = spot_n_488 ))

spot_B1IIIb_violin + geom_violin(fill = "green4") + geom_boxplot(outlier.shape = NA, width=0.03, color="black")+
            ylim(0,25)+
            labs(y = "FGFR2-IIIb HCR Spots per cell") +
                theme_classic()+
                theme(
                legend.position ="none",
                plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
                axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 20, color="black", face="bold"),
                axis.text = element_text(size=20,color="black", face="bold"))
```

Plotting FGFR2-IIIb in B1-IIIB System with p-values

```{r Fig3_B_2,fig.width=10,fig.height=8}

spot_B1IIIb_violin_pv <- ggplot(dt_IIIB, aes(x = cell_line, y = spot_n_488 ))

spot_B1IIIb_violin_pv + geom_violin(fill = "green4") + geom_boxplot(outlier.shape = NA, width=0.03, color="black")+
  
            stat_pvalue_manual(IIIB_pwc_p, hide.ns = TRUE) +
            labs(y = "FGFR2-IIIb HCR Spots per cell") +
                theme_classic()+
                theme(
                legend.position ="none",
                plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
                axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 20, color="black", face="bold"),
                axis.text = element_text(size=20,color="black", face="bold"))
```



Plotting PGK1 in B2-PGK1 System.

```{r Fig3_C,fig.width=10,fig.height=8, echo=FALSE}
spot_B2PGK1_violin <- ggplot(dt_PGK1, aes(x = cell_line, y = spot_n_562))

spot_B2PGK1_violin + geom_violin(fill = "grey70") + geom_boxplot(outlier.shape = NA, width=0.03, color="black")+
            ylim(0,250)+
            labs(y = "PGK1 HCR Spots per cell") +
                theme_classic()+
                theme(
                legend.position ="none",
                plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
                axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 20, color="black", face="bold"),
                axis.text = element_text(size=20,color="black", face="bold"))
```

Plotting PGK1 in B2-PGK1 System with p-values

```{r Fig3_C_2,fig.width=10,fig.height=8, echo=FALSE}

spot_B2PGK1_violin_pv <- ggplot(dt_PGK1, aes(x = cell_line, y = spot_n_562))

spot_B2PGK1_violin_pv + geom_violin(fill = "grey70") + geom_boxplot(outlier.shape = NA, width=0.03, color="black")+
            stat_pvalue_manual(PGK1_pwc_p, hide.ns = TRUE) +
            labs(y = "PGK1 HCR Spots per cell") +
                theme_classic()+
                theme(
                legend.position ="none",
                plot.title = element_text(size = 24, hjust = 0.5, face = "bold"),
                axis.title.x = element_blank(),
                axis.text.x = element_text(angle = 45, hjust = 1),
                axis.title = element_text(size = 20, color="black", face="bold"),
                axis.text = element_text(size=20,color="black", face="bold"))
```



Document the information about the analysis session. 

```{r sessionInfo, include=TRUE, echo=TRUE, results="markup"}
sessionInfo()
```
